---
title: "4840 data cleaning"
author: "Rebecca"
date: "`r Sys.Date()`"
output: html_document
---

Data cleaning 
Input : train.csv
Outputs : 
df            - cleaned data
df_no_out     - cleaned data with no outliers (using boxplot method only) on the 2 Delay.in.Minutes columns
X_scale       - scaled df without target variable (dum_satisfaction)
X_scale_noout - scaled df_no_out without target variable (dum_satisfaction)
y             - target feature column


Cleaning done :
A1 Remove X and id as they are unique identifiers, and extract personal travel records
A2 Missing values of Arrival.Delay.in.Minutes => median for each satisfaction group
A3 zero rating value = 'not applicable' - removed
A4 generate int dummy for categorical data and drop the original columns

Dummy variables for categorical features:
==========================================
[1] dum_Gender : "Female 0" "Male 1"  
[1] dum_Customer : "disloyal Customer 0" "Loyal Customer 1"   
[1] dum_Class : "Business 2" "Eco 0"      "Eco Plus 1"
[1] dum_satisfaction : "neutral or dissatisfied 0" "satisfied 1"    



# 1. Read the data into R:
```{r}
temp <- read.csv("~/GIT/MY_PROJECTS/KMedoid/data/train.csv")
```

# A1 Remove X and id as they are unique identifiers, and extract personal travel records
```{r}
raw_df<-temp[temp$Type.of.Travel=="Personal Travel",-c(1,2)]
#drop type of travel
raw_df<-raw_df[,-4]
# View(raw_df)
```

32249 records

```{r}
df1=raw_df
```


# A2 - Missing values of Arrival.Delay.in.Minutes => median for each satisfaction group
```{r}
m1 <- median(df1[which(df1$satisfaction == "satisfied" & !is.na(df1$Arrival.Delay.in.Minutes)),]$Arrival.Delay.in.Minutes)
m2 <- median(df1[which(df1$satisfaction != "satisfied" & !is.na(df1$Arrival.Delay.in.Minutes)),]$Arrival.Delay.in.Minutes)
df1$Arrival.Delay.in.Minutes <- ifelse(is.na(df1$Arrival.Delay.in.Minutes)&df1$satisfaction=='satisfied', 
                            m1, ifelse(is.na(df1$Arrival.Delay.in.Minutes)&df1$satisfaction!='satisfied', 
                                             m2, df1$Arrival.Delay.in.Minutes))
```


# A3 - zero rating value = 'not applicable' - remove

```{r}
(all_col<-colnames(df1))
```

```{r}
# check percentage of records with ratings = 0 for each feature
n=nrow(df1)
score_col=all_col[6:19]
score_col_impute = c()   # record the columns that may need to be imputed later
for (col in score_col) {
  pc = 100* (nrow(df1[df1[col]==0,]) / n)
  med = median(df1[,col])
  if (pc > 0) { 
    score_col_impute <- c(score_col_impute, col)
    }
  print(paste(col, pc, med))
}


```


```{r}
# columns to impute (values stored during eda)
score_col_impute
```

```{r}
for (col in score_col_impute) {
  df1[,col] <- ifelse(df1[,col]==0 , NA, df1[,col])
}

```

drop all na data
```{r}
df_nomiss <- na.omit(df1)
```

```{r}
summary(df_nomiss)
```

29598 records left

# A4 generate int dummy for categorical data

[1] "Gender c(\"Male\", \"Female\")"
[1] "Customer.Type c(\"Loyal Customer\", \"disloyal Customer\")"
[1] "Class c(\"Eco Plus\", \"Eco\", \"Business\")"
[1] "satisfaction c(\"neutral or dissatisfied\", \"satisfied\")"

```{r}
require(plyr)
g = sort(unique(df_nomiss$Gender))
order=c(0:(length(g)-1))
print(paste(g, order))
df_nomiss$dum_Gender <- mapvalues(df_nomiss$Gender, from=g, to=order)
df_nomiss$dum_Gender <- as.integer(df_nomiss$dum_Gender)

# 
g = sort(unique(df_nomiss$Customer.Type))
order=c(0:(length(g)-1))
print(paste(g, order))
df_nomiss$dum_Customer <- mapvalues(df_nomiss$Customer.Type, from=g, to=order)
df_nomiss$dum_Customer <- as.integer(df_nomiss$dum_Customer)

# "Business 0" "Eco 1"      "Eco Plus 2"
g = sort(unique(df_nomiss$Class))
order=c(2,0,1)
# order=c(0:(length(g)-1))
print(paste(g, order))
df_nomiss$dum_Class <- mapvalues(df_nomiss$Class, from=g, to=order)
df_nomiss$dum_Class <- as.integer(df_nomiss$dum_Class)

# 
g = sort(unique(df_nomiss$satisfaction))
order=c(0:(length(g)-1))
print(paste(g, order))
df_nomiss$dum_satisfaction <- mapvalues(df_nomiss$satisfaction, from=g, to=order)
df_nomiss$dum_satisfaction <- as.integer(df_nomiss$dum_satisfaction)
```

```{r}
# remove original categorical data
df<- df_nomiss[, -c(1,2,4,22)]
```

```{r}
summary(df)
```


```{r}
outliers <- boxplot(df$Departure.Delay.in.Minutes, plot=FALSE)$out
```

```{r}
df_no_out<-df
df_no_out<- df_no_out[-which(df_no_out$Departure.Delay.in.Minutes %in% outliers),]
```

```{r}
outliers2 <- boxplot(df_no_out$Arrival.Delay.in.Minutes, plot=FALSE)$out
```

```{r}
df_no_out<- df_no_out[-which(df_no_out$Arrival.Delay.in.Minutes %in% outliers2),]
```


```{r}
colnames(df_no_out)
```

```{r}
library(stats)
library(psych)
library("corrplot")

visualize <- function(data, numcols) {
  cor.plot(df[,numcols])
  corr = cor(df[,-numcols], method='s')
  corrplot(corr, method = "color",
     type = "upper", order = "hclust", 
     #addCoef.col = "black",
     tl.col = "black")  
}
```

```{r}
numcols=c(1,2,17,18)
visualize(df, numcols)
```

- correlation on servicing features (On.board.service, baggage handling, inflight.services); technology (online.boarding, inflight.wifi.service, Ease.of.online.booking); entertainment (Seat.comfort, inflight.entertainment, food.and.drink, Cleanliness)

```{r}
data0=df[df$dum_satisfaction==0,]
visualize(data0, numcols)
```


```{r}
data1=df[df$dum_satisfaction==1,]
visualize(data1, numcols)
```
missing values count by column
```{r}
library(dlookr)
diagnose(df)
```


Scale data (df with outliers)
```{r}
target_col=ncol(df)
X = df[,-target_col]
y = df[,target_col]

#scale the data
X_scale=as.data.frame(scale(X))
cnames=colnames(X)
scale_tbl=data.frame()
# store the mean and std for each column for future use if needed
for (i in c(1:21)) {
  m = mean(X[,i])
  s = sd(X[,i])
  scale_tbl =rbind(scale_tbl, c(cnames[i],m,s))
}

colnames(scale_tbl) <- c("var.names","mean", "sdev")
# scale_tbl
```


Scale data (df with no outliers)
```{r}
target_col=ncol(df)
X_noout = df[,-target_col]

#scale the data
X_scale_noout=as.data.frame(scale(X_noout))
cnames=colnames(X_noout)
scale_tbl_noout=data.frame()
# store the mean and std for each column for future use if needed
for (i in c(1:21)) {
  m = mean(X_noout[,i])
  s = sd(X_noout[,i])
  scale_tbl_noout =rbind(scale_tbl_noout, c(cnames[i],m,s))
}

colnames(scale_tbl_noout) <- c("var.names","mean", "sdev")
# scale_tbl_noout
```